A generalization of Plackett-Luce likelihoods to incorporate non-independence of competitors

### 2024-01-26 inst/NBA.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "An analysis of a single basketball game"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I consider a single basketball match from the perspective of the
Bradley-Terry model.  File `NBA.txt` contains a point-by-point
analysis of a basketball match between the Cleveland Cavaliers and the
Golden State Warriors on 13 June 2017, data taken from
[here](https://www.espn.com/nba/playbyplay?gameId=400954514).  Some
further documentation is given at `inst/NBA.Rd` in the package.

>>>>

---
title: "An analysis of a single basketball game"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here I consider a single basketball match from the perspective of the
Bradley-Terry model.  File `NBA.txt` contains a point-by-point
analysis of a basketball match between the Cleveland Cavaliers and the
Golden State Warriors on 13 June 2017, data taken from
[here](https://www.espn.com/nba/playbyplay?gameId=400954514).  Some
further documentation is given at `inst/NBA.Rd` in the package.

### 2024-01-26 inst/T20.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "The impact of winning the coin toss on the Indian Premier League"
author: "Tanvi Chandel and Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This script creates and analyses hyper2 object `T20` which is a
likelihood function for the strengths of the competitors in the Indian
Premier League.
Dataframe `T20_table` has one row for each T20 IPL match in the
period 2008-2017 with the exception of three no-result matches and
seven tied matches, which were removed.

```{r kickoff}
library("hyper2",quietly=TRUE)
T20_table <- read.table("T20.txt",header=TRUE)
head(T20_table)
nrow(T20_table)
```

>>>>

---
title: "The impact of winning the coin toss on the Indian Premier League"
author: "Tanvi Chandel and Robin K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy hyper2.bib from the vignettes/ directory
---

@@

To cite the `hyper2` package in publications, please use
@hankin2017_rmd.  This script creates and analyses hyper2 object `T20`
which is a likelihood function for the strengths of the competitors in
the Indian Premier League.  Dataframe `T20_table` has one row for each
T20 IPL match in the period 2008-2017 with the exception of three
no-result matches and seven tied matches, which were removed.

```{r kickoff}
library("hyper2",quietly=TRUE)
library("magrittr",quietly=TRUE)
T20_table <- read.table("T20.txt",header=TRUE)
head(T20_table)
nrow(T20_table)
```

### 2024-01-26 inst/carcinoma.Rmd
distinguish hankin2017 from hankin2017_rnw

To cite the `hyper2` package in publications, please use @hankin2017
(takes maybe twenty minutes to run without cache.  The value of
`carcinoma_maxp` is calculated separately [discussed below], and that
takes a few days to run).

>>>>

To cite the `hyper2` package in publications, please use @hankin2017_rmd; 
takes maybe twenty minutes to run without cache.  The value of
`carcinoma_maxp` is calculated separately [discussed below], and that
takes a few days to run.

### 2024-01-26 inst/collusion3.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Chess collusion with hyper3 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I analyse results from Curacao 1962 using `hyper3` formalism
using the weighted Draw monster approach outlined in `kka.Rmd`:

>>>>

---
title: "Chess collusion with hyper3 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy from vignettes/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here I analyse results from Curacao 1962 using `hyper3` formalism
using the weighted Draw monster approach outlined in `kka.Rmd`:

### 2024-01-26 inst/cook_repeated.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Judging domestic cooking ability: repeated incomplete rank observations"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017;
takes about five minutes without cache.  Suppose there are five people
in a family.  Each day for five days, one person cooks a meal.  After
the fifth day, each person considers the four meals cooked by other
family members, and puts them in order of preference.  This process is
repeated and here I consider how to interpret such data.

>>>>

---
title: "Judging domestic cooking ability: repeated incomplete rank observations"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib # copy from vignettes/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
takes about five minutes without cache.  Suppose there are five people
in a family.  Each day for five days, one person cooks a meal.  After
the fifth day, each person considers the four meals cooked by other
family members, and puts them in order of preference.  This process is
repeated and here I consider how to interpret such data.

### 2024-01-26 inst/curling.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "curling with hyper2 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
[Curling](https://en.wikipedia.org/wiki/Curling) is a sport with
stones and brooms.  Following dataset taken from
[wikipedia](https://en.wikipedia.org/w/index.php?title=Curling_at_the_Winter_Olympics&oldid=998891075),
women's tournament:

>>>>

---
title: "curling with hyper2 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy from vignettes/

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
[Curling](https://en.wikipedia.org/wiki/Curling) is a sport with
stones and brooms.  Following dataset taken from
[wikipedia](https://en.wikipedia.org/w/index.php?title=Curling_at_the_Winter_Olympics&oldid=998891075),
women's tournament:

### 2024-01-26 inst/distance_matrix_idea.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "icons: some experimental results"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This file presents some _speculative_ experimental work which is under
development.  Do not confuse this file with `icons.Rmd`, which creates
the `icons` objects in the package.

>>>>

---
title: "icons: some experimental results"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib # copy from vignettes/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This file presents some _speculative_ experimental work which is under
development.  Do not confuse this file with `icons.Rmd`, which creates
the `icons` objects in the package.

### 2024-01-26 inst/eurodance.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Eurovision **Dance** Contest 2008"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017;
takes about 20 minutes to process without cache.  The file
`eurodance2008.txt`, used below, is copied from "Eurovision Dance
Contest 2008", Wikipedia, accessed March 31, 2021.  First we specify
the matrix as appearing in the Wikipedia page:

>>>>

---
title: "Eurovision **Dance** Contest 2008"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy from vignettes/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
takes about 20 minutes to process without cache.  The file
`eurodance2008.txt`, used below, is copied from "Eurovision Dance
Contest 2008", Wikipedia, accessed March 31, 2021.  First we specify
the matrix as appearing in the Wikipedia page:

### 2024-01-26 inst/fatigue.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "fatigue"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Analysis of the 1973 interzonal competition.

>>>>

---
title: "fatigue"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Analysis of the 1973 interzonal competition.

### 2024-01-26 inst/global_liveability_ranking.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Global liveability ranking"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017;
takes about an hour to process without cache.

>>>>

---
title: "Global liveability ranking"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
takes about an hour to process without cache.

### 2024-01-26 inst/grad_weight.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Differentiation with respect to weights for a `hyper3` object"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
We will start with a simple example.  Consider two players $p_1$ and
$p_2$ who play tennis say, $a+b$ times with $p_1$ winning $a$ times
and $p_2$ winning $b$ times.  We use the following likelihood function:

>>>>

---
title: "Differentiation with respect to weights for a `hyper3` object"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
We will start with a simple example.  Consider two players $p_1$ and
$p_2$ who play tennis say, $a+b$ times with $p_1$ winning $a$ times
and $p_2$ winning $b$ times.  We use the following likelihood function:

### 2024-01-26 inst/handover.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "handover issues in medical care"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017.
This short document is based on @altham2010, which
discussed an interesting dataset originally presented by Lin
et. al. (2009).  It uses more sophisticated idiom furnished by the
`hyper2` package, which is an improved and efficient upgrade of the
primitive `hyperdirichlet` package.  The dataset arose from 69 medical
malpractice claims and refers to the two surgeon–reviewers' answers to
the question "was there a communication breakdown in the hand-off
between physicians caring for the patient?".  The rows of Table 1
correspond to the answers that were given by reviewer 1, and the
columns to the answers given by reviewer 2.

>>>>

---
title: "handover issues in medical care"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This short document is based on @altham2010, which
discussed an interesting dataset originally presented by Lin
et. al. (2009).  It uses more sophisticated idiom furnished by the
`hyper2` package, which is an improved and efficient upgrade of the
primitive `hyperdirichlet` package.  The dataset arose from 69 medical
malpractice claims and refers to the two surgeon–reviewers' answers to
the question "was there a communication breakdown in the hand-off
between physicians caring for the patient?".  The rows of Table 1
correspond to the answers that were given by reviewer 1, and the
columns to the answers given by reviewer 2.

### 2024-01-26 inst/hepatitis.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Hepatitis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.

>>>>

---
title: "Hepatitis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.

### 2024-01-26 inst/home_advantage.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Quantifying the home ground advantage with `hyper2` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017
(takes about fifteen minutes to run without cache).  Here I analyse a
small synthetic dataset for home ground advantage, then go on to
discuss a dataset presented by Agresti.  But first, the smallest
nontrivial example is a 2x2 case:

>>>>

---
title: "Quantifying the home ground advantage with `hyper2` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd
(takes about fifteen minutes to run without cache).  Here I analyse a
small synthetic dataset for home ground advantage, then go on to
discuss a dataset presented by Agresti.  But first, the smallest
nontrivial example is a 2x2 case:

### 2024-01-26 inst/home_advantage3.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Quantifying the home advantage with `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

>>>>

---
title: "Quantifying the home advantage with `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Objects of class \code{hyper3} are a generalization of \code{hyper2}
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

### 2024-01-26 inst/hunter.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Iteration techniques for Plackett-Luce likelihood functions and the `hyper2` package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I discuss why implementing equation 30 of [@hunter2004] for
`hyper2` objects representing Plackett-Luce likelihood functions is
difficult.  Consider the following rank observations: `724356` [that
is, 7 came first, then 2, then 4 then 3 then 5 then 6] and `1372`
[that is, 1 came first, then 3 then 7 then 2].  If these two
observations are independent, we would have a Plackett-Luce likelihood
function of

>>>>

---
title: "Iteration techniques for Plackett-Luce likelihood functions and the `hyper2` package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here I discuss why implementing equation 30 of [@hunter2004] for
`hyper2` objects representing Plackett-Luce likelihood functions is
difficult.  Consider the following rank observations: `724356` [that
is, 7 came first, then 2, then 4 then 3 then 5 then 6] and `1372`
[that is, 1 came first, then 3 then 7 then 2].  If these two
observations are independent, we would have a Plackett-Luce likelihood
function of

### 2024-01-26 inst/hyper3.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "The weighted hyperdirichlet distribution: the `hyper3` class"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017.
Objects of class `hyper3` are a generalization of `hyper2`
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative numbers $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

>>>>

---
title: "The weighted hyperdirichlet distribution: the `hyper3` class"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Objects of class `hyper3` are a generalization of `hyper2`
objects that allow the brackets to contain weighted probabilities.
Likelihood functions are defined on non-negative numbers $p_1,\ldots, p_n$
subject to the unit-sum constraint $\sum p_i=1$.  Given known weights
$w^i_j$ with $1\leq i\leq j$ we have

### 2024-01-26 inst/icons.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Analysing icons for climate change and with `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This minimal document reads file `icons.txt` and creates R objects
`icons`, `icons_matrix`, and `icons_maxp`.  The dataset is discussed
extensively in @west2008 and @hankin2010.

>>>>

---
title: "Analysing icons for climate change and with `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This minimal document reads file `icons.txt` and creates R objects
`icons`, `icons_matrix`, and `icons_maxp`.  The dataset is discussed
extensively in @west2008 and @hankin2010.

### 2024-01-26 inst/interzonal.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Analysis of the 1962 world chess championship, Curacao"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This file creates likelihood functions `H` and `H_coll`, which appear
in the package as `interzonal` and `collusion` respectively.  It also
presents some analysis of the datasets using the `hyper2` package.

>>>>

---
title: "Analysis of the 1962 world chess championship, Curacao"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This file creates likelihood functions `H` and `H_coll`, which appear
in the package as `interzonal` and `collusion` respectively.  It also
presents some analysis of the datasets using the `hyper2` package.

### 2024-01-26 inst/javelin.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Javelins with hyper3 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here we analyse results from the men's javelin, 2020 Summer Olympics.

>>>>

---
title: "Javelins with hyper3 objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here we analyse results from the men's javelin, 2020 Summer Olympics.

### 2024-01-26 inst/jester.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "The Jester joke recommender system"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
This file creates objects `jester` and `maxjest`, which are datasets
available in the `hyper2` package, and documented at `jester.Rd`.
This file takes quite a long time to run.  Goldberg et al present a
dataset in which respondents rated a number of jokes.  Here, I analyse
a small portion of this dataset using the `hyper2` package.  This
document is intended to illustrate an extremely challenging
application of the `hyper2` package and (without cache) takes a long
time to process.  Goldberg's dataset has 24938 lines, one per
respondent, and 101 columns, one per joke (the first column shows the
number of jokes rated by each respondent); here I use 150 lines and 99
jokes (the 100th joke was not funny).

>>>>

---
title: "The Jester joke recommender system"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
This file creates objects `jester` and `maxjest`, which are datasets
available in the `hyper2` package, and documented at `jester.Rd`.
This file takes quite a long time to run.  Goldberg et al present a
dataset in which respondents rated a number of jokes.  Here, I analyse
a small portion of this dataset using the `hyper2` package.  This
document is intended to illustrate an extremely challenging
application of the `hyper2` package and (without cache) takes a long
time to process.  Goldberg's dataset has 24938 lines, one per
respondent, and 101 columns, one per joke (the first column shows the
number of jokes rated by each respondent); here I use 150 lines and 99
jokes (the 100th joke was not funny).

### 2024-01-26 inst/karate.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "karate"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I analyse a [karate](https://en.wikipedia.org/wiki/Karate)
dataset taken from
[wikipedia](https://en.wikipedia.org/wiki/2018_World_Karate_Championships).
Short story: the dataset has too many degrees of freedom to be easily
analysed with `hyper2` and Zermelo's iteration scheme as implemented
by the `BradleyTerry` package exploits the special pairwise structure
of the likelihood function very effectively.

>>>>

---
title: "karate"
author: "R. K. S. Hankin"
output: bookdown::html_document2  
bibliography: hyper2.bib # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here I analyse a [karate](https://en.wikipedia.org/wiki/Karate)
dataset taken from
[wikipedia](https://en.wikipedia.org/wiki/2018_World_Karate_Championships).
Short story: the dataset has too many degrees of freedom to be easily
analysed with `hyper2` and Zermelo's iteration scheme as implemented
by the `BradleyTerry` package exploits the special pairwise structure
of the likelihood function very effectively.

### 2024-01-26 inst/kka.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "A Bradley-Terry analysis of three grandmasters: Karpov, Kasparov, Anand"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017;
this `Rmd` file takes about 45 minutes to process without cache.  Here
we define dataset `kka` which is the observed scorelines between
players Karpov, Kasparov, and Anand.  It is used to calculate
likelihood functions: `karpov_kasparov_anand`, `kka_3whites` and
`kka_3draws`.  This Rmd file supercedes files
`inst/karpov_kasparov_anand.R` and `inst/kka_3draws.R` and
`inst/kka_array.R`.  Some more detail is given in `man/kka.Rd`.

>>>>

---
title: "A Bradley-Terry analysis of three grandmasters: Karpov, Kasparov, Anand"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
this `Rmd` file takes about 45 minutes to process without cache.  Here
we define dataset `kka` which is the observed scorelines between
players Karpov, Kasparov, and Anand.  It is used to calculate
likelihood functions: `karpov_kasparov_anand`, `kka_3whites` and
`kka_3draws`.  This Rmd file supercedes files
`inst/karpov_kasparov_anand.R` and `inst/kka_3draws.R` and
`inst/kka_array.R`.  Some more detail is given in `man/kka.Rd`.

### 2024-01-26 inst/loser.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Weakly informative observations: identifying the weakest player"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017;
see also `inst/notthewinner.Rmd`.  Here we consider a situation in
which we make only a rather uninformative observation: namely, the
identity of the _loser_ (which is here taken to be me, player `a`).
We have six competitors `a,b,c,d,e,f` with nonnegative strengths
$p_1,p_2,p_3,p_4,p_5,p_6, \sum p_i=1$.  Competitor `a` has strength
$p_1$, and loses.  We are interested in how informative this
observation is with regard to various assertions about the strength of
`a`.  We have three hypotheses:

>>>>

---
title: "Weakly informative observations: identifying the weakest player"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
see also `inst/notthewinner.Rmd`.  Here we consider a situation in
which we make only a rather uninformative observation: namely, the
identity of the _loser_ (which is here taken to be me, player `a`).
We have six competitors `a,b,c,d,e,f` with nonnegative strengths
$p_1,p_2,p_3,p_4,p_5,p_6, \sum p_i=1$.  Competitor `a` has strength
$p_1$, and loses.  We are interested in how informative this
observation is with regard to various assertions about the strength of
`a`.  We have three hypotheses:

### 2024-01-26 inst/mann_whitney_wilcoxon.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Generalizing the Mann-Whitney-Wilcoxon test with `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I discuss the `hyper2` generalization of the
Mann-Whitney-Wilcoxon test, applying it to the chorioamnion dataset at
`wilcox.test.Rd`:

>>>>

---
title: "Generalizing the Mann-Whitney-Wilcoxon test with `hyper3` objects"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here I discuss the `hyper2` generalization of the
Mann-Whitney-Wilcoxon test, applying it to the chorioamnion dataset at
`wilcox.test.Rd`:

### 2024-01-26 inst/monster_vs_lambda.Rmd
distinguish hankin2017 from hankin2017_rnw

---
title: "Monster or Lambda?  A likelihood analysis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017;
takes about an hour to process without cache.  Here I consider
pairwise comparisons with a home-ground advantage and wish to compare
two likelihood functions.  First, the monster parametrization:

>>>>

---
title: "Monster or Lambda?  A likelihood analysis"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd;
takes about an hour to process without cache.  Here I consider
pairwise comparisons with a home-ground advantage and wish to compare
two likelihood functions.  First, the monster parametrization:

### 2024-01-26 inst/multiset_hyper2.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "The `cheering3()` probability model for order statistics: a
verification with the `multiset()` function of the `partitions` package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here we will sum probabilities across a sample space and verify that
the total is constant.  In theory the sum of the probabilities of all
members of the sample space is 1, but in practice [and certainly in
`hyper2` idiom], the likelihood calculation includes an indeterminate
constant which means that all we can say is that the total probability
is not a function of the parameters of the distribution.

>>>>

---
title: "The `cheering3()` probability model for order statistics: a
verification with the `multiset()` function of the `partitions` package"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here we will sum probabilities across a sample space and verify that
the total is constant.  In theory the sum of the probabilities of all
members of the sample space is 1, but in practice [and certainly in
`hyper2` idiom], the likelihood calculation includes an indeterminate
constant which means that all we can say is that the total probability
is not a function of the parameters of the distribution.

### 2024-01-26 inst/order.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Order statistics and numerical optimization: the case of a single observation"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I consider the likelihood function given by a single observation
of an order statistic, which is problematic in the sense that the
maximum likelihood estimate for the BT strengths of the competitors is
$(1,0,\ldots,0)$.  Here we investigate some numerical problems arising
from the use of `optim()` and `constrOptim()`.  We will consider
$n=10$ competitors and without loss of generality, the competitors
finish in order $1,2,3,\ldots,10$.  The likelihood function for this
observation is:

>>>>

---
title: "Order statistics and numerical optimization: the case of a single observation"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here I consider the likelihood function given by a single observation
of an order statistic, which is problematic in the sense that the
maximum likelihood estimate for the BT strengths of the competitors is
$(1,0,\ldots,0)$.  Here we investigate some numerical problems arising
from the use of `optim()` and `constrOptim()`.  We will consider
$n=10$ competitors and without loss of generality, the competitors
finish in order $1,2,3,\ldots,10$.  The likelihood function for this
observation is:

### 2024-01-26 inst/ordertable_to_ranktable.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "The print method for class ranktable"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Rank tables can be confusing.  The package includes a range of
functionality for dealing with them, including a bespoke print method.
Here, I discuss two cases: firstly, a tiny toy example, and secondly
the volvo ocean racing dataset.

>>>>

---
title: "The print method for class ranktable"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Rank tables can be confusing.  The package includes a range of
functionality for dealing with them, including a bespoke print method.
Here, I discuss two cases: firstly, a tiny toy example, and secondly
the volvo ocean racing dataset.

### 2024-01-26 inst/pentathlon.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "Pentathlon"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
The [modern
pentathlon](https://en.wikipedia.org/wiki/Modern_pentathlon) is an
Olympic sport that comprises five different events: fencing, swimming,
horse riding, and running.  This short document discusses a dataset
taken from the [2004 Summer
Olympics](https://en.wikipedia.org/wiki/Modern_pentathlon_at_the_2004_Summer_Olympics),
analysing the order statistic for the top seven competitors:

>>>>

---
title: "Pentathlon"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
The [modern
pentathlon](https://en.wikipedia.org/wiki/Modern_pentathlon) is an
Olympic sport that comprises five different events: fencing, swimming,
horse riding, and running.  This short document discusses a dataset
taken from the [2004 Summer
Olympics](https://en.wikipedia.org/wiki/Modern_pentathlon_at_the_2004_Summer_Olympics),
analysing the order statistic for the top seven competitors:

### 2024-01-26 inst/plackett_luce_monster.Rmd
consistentification
distinguish hankin2017 from hankin2017_rnw

---
title: "A generalization of Plackett-Luce likelihoods to incorporate non-independence of competitors"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: jss_hyper3.bib
---

@@

To cite the `hyper2` package in publications, please use @hankin2017a.
Here I modify standard Plackett-Luce likelihood functions to include
an element of non-independence among the competitors.  There seem to
be a few distinct statistical models for nonindependence:

>>>>

---
title: "A generalization of Plackett-Luce likelihoods to incorporate non-independence of competitors"
author: "R. K. S. Hankin"
output: bookdown::html_document2
bibliography: hyper2.bib  # copy vignettes/hyper2.bib to inst/
---

@@

To cite the `hyper2` package in publications, please use @hankin2017_rmd.
Here I modify standard Plackett-Luce likelihood functions to include
an element of non-independence among the competitors.  There seem to
be a few distinct statistical models for nonindependence:
